\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-12 fully-abstract-comp/popl2019/paper.tex
Small changes to section 5.

\input{../preamble}
% Paper specific redefinitions of commands
\renewcommand{\MemSeg}{\shareddom{MemSeg}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\renewcommand{\perm}{\var{p}}
\renewcommand{\SealableCaps}{\shareddom{Sealables}}
\renewcommand{\Cap}{\shareddom{NatTok}}
\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
\renewcommand{\comp}{\var{comp}}
\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
\renewcommand{\Worlds}{\World_{\mathrm{call\ stack}}}

@@

\begin{abstract}
  We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
\end{abstract}

@@

%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
In this paper, we make the following contributions:
% Presents the first formal model of a capability machine with linear capabilities (Section 2)
% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
\begin{itemize}
\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
\item We present a new way to formalise these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
  \begin{itemize}
  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
  \end{itemize}
\end{itemize}
This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.

@@

%%% Capabilities
The concept of a capability is the cornerstone of any capability machine.
In its simplest form, a capability is a permission and a range of authority.
The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
% Permissions and Range of authority
The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}). Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$), ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
% Pointer and linearity
In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
% What is linear capabilities, split/splice
A linear capability is a capability that cannot be duplicated.
This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.

@@

\subsection{Logical relation}
\label{subsec:logical-relation}
In this section, we present the logical relation.
Unfortunately, for space reasons, we do not give a comprehensive description of it.
Instead we highlight the parts specific to the logical relation and refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
Finally, for presentation, we will sometimes omit details, like step indexing, that are important for correctness but otherwise not interesting.

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, a world is used in a logical relation to model assumptions about memory.
\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from $\RegionName$ (modelled as natural numbers) to regions.
The regions differ for each of the sub-worlds based on what they model.
Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
The region types model linearity as they indicate how the memory can be referenced.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be referenced by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
To this end, we have the partial operation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except that they have disjoint ownership: $\pure$ regions can be present in both worlds, but a $\spatialo$ region can only be present in one.
Such a region must then be $\spatial$ in the other world.
%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
This is important because the logical relation will require that valid memories are backed by a suitable memory segment, even when the memory does not own them (for example, when a linear capability for it is in one of the registers).
Revocation of linear capabilities is supported in a certain way using the $\revoked$ region.
Finally, $\pure$ regions can also claim ownership of a seal by defining a seal interpretation for it.
The seal interpretation is a world-indexed relation that must hold for all sealables signed by this seal.
We split the regions in two categories:
\begin{align*}
  \Regionh &= 
             \left\{\begin{multlined}
               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
             \end{multlined}\right.\\
  \Regions &={}
             \left\{\begin{multlined}
               \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
               \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
               \{\revoked\}
             \end{multlined}\right.
\end{align*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
The call stack and free stack are only ever addressed by linear capabilities, so their sub-worlds only allow $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they contain the return address for this frame.

Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$. They are recursive as they contain world indexed relations. Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct $\Wor$ which solves the circularity\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
% \begin{lemma}
%   \label{thm:recursive-domain-eq}
%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
%   \[
%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
%   \]
%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
% \end{lemma}
The future world relation $\future$ is fairly standard.
It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
$\spatial$ regions can become $\spatialo$ or $\revoked$.

@@

The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigclos$ can seal.
The return seals can only be used to seal return pointers that corresponds to the return from a call in the code of the region.
Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
The code return pointer corresponds to a normal capability which means that it can be reused but always to return from the same call.
Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function relates data return pointers to linear capabilities.
It also requires the data return pointer to correspond to a $\spatialo$ region for one of the stack frames on the call stack.
The return address that is part of this region in $\Worlds$, must be the return address of the call that the return seal is affiliated with.
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\mathrm{xjmp\_off}+1)\right)} \mid \dots
    %     \sigrets \subseteq \gsigrets \tand \\
    %     \quad\dom(\mscode) \subseteq \ta \tand 
    %     \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
    %     \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
    %     \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
    %     \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
    \end{array}
      \right\} \cup \\
\left\{
    \begin{array}{l}
      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid \dots
        % \sigrets \subseteq \gsigrets \tand 
        % \dom(\mscode) \subseteq \ta \tand\\
        % \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
        % \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        % \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        % \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        % \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets
\end{multline*}

Unlike return seals, closure seals are allowed in both trusted and untrusted components, but we cannot allow them to use closure seals to seal the same things.
If we allowed this, then a trusted component could seal a read-only capability for its code (i.e.\ access to return seals) with an untrusted closure seal and give it to the untrusted component which can unseal it with \texttt{xjmp}.
The closure seals in untrusted code should therefore only be used to seal things from the untrusted value relation, i.e.\ things that are safe to transfer to an untrusted component.
The trusted closure seals can by used to seal things from the untrusted value relation.
This allows trusted seals to seal capabilities for the trusted code which is necessary in order to create a new closure.

@@

Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, the they cannot depend on the same spatially owned region.
% memory relation
The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that corresponds with the view \srccm{} provides on the memory.
Each of the three partitions of $\ms_T$ must be related to the three parts of \srccm{} memory under the relevant sub-worlds of $W$.
The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
The free stack $\ms_\stk$ relates to the free stack partition of $\ms_T$ in the standard way, but with respect to the free stack world.
%Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts relating each of the stack frames.
Further, as each of the stack frames correspond to a slot in the return order, none of the stack frames should be empty.
It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved in the actual memory.

@@

\subsubsection{Value relations}
The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonable, e.g.\ by treating return seals in a reasonable way.
This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and that we want to guarantee LSE and WBCF against.
This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and untrusted value relation.
The two value relations are sketched in Figure~\ref{fig:value-relation}.

The untrusted value relation relates words that can safely be given to untrusted code.
Data is always safe, and a piece of data relates to itself.
A stack pointer is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
The stack pointer should only be able to produce safe words, so a stack pointer with read permission can only produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a stack pointer with write permission can at least write safe words.
Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack part of the world.
Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read-write permission.
The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
An executable memory capability should always point to code which means that there should be a code region governing the memory it points to.
The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seals is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
% The write condition is called $\stackWriteCond{}$
% \begin{multline*}
%   \stackWriteCond{W} =\\
%   \left\{
%     \begin{array}{l}\npair{A} \mid 
%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
%     \end{array}
%   \right\}
% \end{multline*}
When a piece of untrusted code is called, it is passed a return pair consisting of two sealed capabilities.
Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\vsc_S,\vsc_T)$ must be in the seal interpretation function for $\sigma$.
Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.

@@

The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward.

@@

    $\dom(\comp.\mscode) \subseteq \ta$ and $\comp$ is a reasonable component; or

>>>>

\input{../preamble}
% Paper specific redefinitions of commands
\renewcommand{\MemSeg}{\shareddom{MemSeg}}
\renewcommand{\Reg}{\shareddom{RegFile}}
\renewcommand{\RegName}{\shareddom{RegName}}
\renewcommand{\decInstr}[1]{\plainfun{decode}{#1}}
\renewcommand{\updPcAddr}[1]{\plainfun{updPc}{#1}}
\renewcommand{\linCons}[1]{\plainfun{linClear}{#1}}
\renewcommand{\nonExec}[1]{\plainfun{nonExec}{#1}}
\renewcommand{\perm}{\var{p}}
\renewcommand{\SealableCaps}{\shareddom{Sealables}}
\renewcommand{\Cap}{\shareddom{NatTok}}
\renewcommand{\rretc}{\mathrm{r}_{\mathrm{rcode}}}
\renewcommand{\rretd}{\mathrm{r}_{\mathrm{rdata}}}
\renewcommand{\comp}{\var{comp}}
\renewcommand{\memSatStack}[3][n]{\npair[#1]{#2} \in \typesetlr{S}(#3)}
\renewcommand{\memSatFStack}[3][n]{\npair[#1]{#2} \in \typesetlt{F}(#3)}
\renewcommand{\Worlds}{\World_{\mathrm{call\_stack}}}
\renewcommand{\Worldfs}{\World_{\mathrm{free\_stack}}}

@@

\begin{abstract}
  We propose and study \stktokens{}: a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
  The calling convention is based on linear capabilities: a type of capabilities that are prevented from being duplicated by the hardware.
  In addition to designing and formalizing this new calling convention, we also contribute a new way to formalize and prove that it effectively enforces well-bracketed control flow and local state encapsulation using what we call a fully abstract overlay semantics.
\end{abstract}

@@

%The main contributions of this paper are (1) \stktokens{} a calling convention that uses linear capabilities to provably secure LSE and WBCF, and (2) a novel proof technique that we call fully-abstract overlay semantics which we apply to prove the properties of \stktokens{}.
In this paper, we make the following contributions:
% Presents the first formal model of a capability machine with linear capabilities (Section 2)
% Presents a calling convention called \stktokens{} that provably guarantee local state encapsulation and well-bracketed control flow (Section 3)
% Presents a novel proof technique that we call fully-abstract overlay semantics. (Section 4-5)
\begin{itemize}
\item We present \trgcm{}: A formalization of a simple CHERI-like capability machine with linear capabilities (Section~\ref{sec:cap-mach-w-seal-and-lin}).
\item We present a new calling convention \stktokens{} that provably guarantees LSE and WBCF on \trgcm{} (Section~\ref{sec:stktokens-explained}).
\item We present a new way to formalize these guarantees based on a novel technique called \textit{fully-abstract overlay semantics} and we prove the results. This includes:
  \begin{itemize}
  \item defining \srccm{}: an overlay semantics for \trgcm{} with built-in LSE and WBCF (Section~\ref{sec:form-secur-with})
  \item proving full-abstraction for a compiler from \srccm{} to \trgcm{} (Section~\ref{sec:fa-proof}) by
  \item using and defining a cross-language, step-indexed, Kripke logical relation with recursive worlds (Section~\ref{sec:fa-proof}).
  \end{itemize}
\end{itemize}
This paper is accompanied by a technical report~\citep{technical_report} with the elided details and proof.

@@

%%% Capabilities
The concept of a capability is the cornerstone of any capability machine.
In its simplest form, a capability is a permission and a range of authority.
The permission dictates the operations the capability can be used for, and the range of authority specifies the range of memory it can act upon.
% Permissions and Range of authority
The capabilities on \trgcm{} are of the form $((\var{perm},\var{lin}),\var{base},\var{end},\var{addr})$ (defined in Figure~\ref{fig:target-syntax} with the rest of the syntax of \trgcm{}). Here $\var{perm}$ is the permission, and $[\var{base},\var{end}]$ is the range of authority.
The available permissions are read-write-execute ($\rwx$), read-write ($\rw$), read-execute ($\rx$), read-only ($\ro$), and null-permission ($\noperm$) ordered by $\le$ as illustrated in Figure~\ref{fig:perm-hier}.
% Pointer and linearity
In addition to the permission and range, capabilities also have a current address $\var{addr}$ and a linearity $\var{lin}$.
The linearity is either $\normal$ for traditional capabilities or $\linear$ for linear ones.
% What is linear capabilities, split/splice
A linear capability is a capability that cannot be duplicated.
This is enforced dynamically on the capability machine, so when a linear capability is moved between registers or memory, the source is cleared.
The non-duplicability of linear capabilities means that a linear capability cannot become aliased if it wasn't to begin with.

@@

\subsection{Logical relation}
\label{subsec:logical-relation}
In this section, we present the logical relation.
Unfortunately, for space reasons, we do not give a comprehensive description of it.
Instead we highlight specific parts to the logical relation and refer to~\citet{technical_report} for details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
Finally, for presentation, we will sometimes omit details, like step indexing, that are important for correctness but otherwise not interesting.

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, a world is used in a logical relation to model assumptions about memory.
\srccm{} has three kinds of memory: the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from $\RegionName$ (modelled as natural numbers) to regions.
The regions differ for each of the sub-worlds based on what they model.
Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
The region types model linearity as they indicate how the memory can be referenced.
We have four types of regions: $\pure$, $\spatial$, $\spatialo$, and $\revoked$.
The $\pure$ regions can be referenced by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ and $\revoked$ regions by nothing.
To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
To this end, we have the partial operation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except that they have disjoint ownership: $\pure$ regions can be present in both worlds, but a $\spatialo$ region can only be present in one.
Such a region must then be $\spatial$ in the other world.
%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
This is important because the logical relation will require that valid memories are backed by a suitable memory segment even when the memory does not own them (for example, when a linear capability for it is in one of the registers).
Revocation of linear capabilities is supported in a certain way using the $\revoked$ region.
Finally, $\pure$ regions can also claim ownership of a seal by defining a seal interpretation for it.
The seal interpretation is a world-indexed relation that must hold for all sealables signed by this seal.
We split the regions in two categories:
\begin{align*}
  \Regionh &= 
             \left\{\begin{multlined}
               \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
               (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
             \end{multlined}\right.\\
  \Regions &={}
             \left\{\begin{multlined}
               \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
               \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
               \{\revoked\}
             \end{multlined}\right.
\end{align*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
The call stack and free stack are only ever addressed by linear capabilities, so their sub-worlds only allow $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they contain the return address for this frame.

Define worlds as $\World = \Worldh \times \Worlds \times \Worldfs$. They are recursive as they contain world indexed relations. Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct $\Wor$ which solves the circularity\footnote{We have cheated in the presentation and included the solution to the circularity already.}.
% \begin{lemma}
%   \label{thm:recursive-domain-eq}
%   There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
%   \[
%     \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
%   \]
%   and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
% \end{lemma}
The future world relation $\future$ is fairly standard.
It is extensional and requires all but $\spatial$ regions to stay the same type in future worlds.
The $\spatial$ regions can become $\spatialo$ or $\revoked$.

@@

The seal interpretation function $H^\mathrm{code,\square}_\sigma \; (\dots)$ dictates what the return seals $\sigrets$ and closure seals $\sigcloss$ can seal.
The return seals can only be used to seal return pointers that corresponds to the return from a call in the code of the region.
Code return pointers must point to the first address after a call, and the call must be associated with the return seal in question.
The corresponding capability on \trgcm{} should point to the address just after the \texttt{xjmp} in the call code (Figure~\ref{fig:call-code}) as it needs to perform the safety checks that are built into the call on \srccm{}.
The code return pointer corresponds to a normal capability which means that it can be reused but only to return from the same call.
Because of this, the code pointer is not associated with any specific call frame as it can be used by any data return pointer generated from the same call point.
Reuse of return pairs is prevented by the data return pointer being linear, so the seal interpretation function relates data return pointers to linear capabilities.
It also requires the data return pointer to refer to a $\spatialo$ region for one of the stack frames on the call stack.
The return address of this region in $\Worlds$ must be the return address of the call that the return seal is affiliated with.
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \mscode \; \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
      \npair{\left(\retptrc(\baddr,\eaddr,\aaddr+\calllen),((\rx,\normal),\baddr,\eaddr,\aaddr+\mathrm{xjmp\_off}+1)\right)} \mid \dots
    %     \sigrets \subseteq \gsigrets \tand \\
    %     \quad\dom(\mscode) \subseteq \ta \tand 
    %     \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
    %     \quad\aaddr = \aaddr' + \retoffset \tand \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand\\
    %     \quad \sigma = \sigma_b + \offsigma \in \sigrets \tand
    %     \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
    \end{array}
      \right\} \cup \\
\left\{
    \begin{array}{l}
      \npair{(\retptrd(\baddr,\eaddr),((\rw,\linear),\baddr,\eaddr,\baddr-1))} \mid \dots
        % \sigrets \subseteq \gsigrets \tand 
        % \dom(\mscode) \subseteq \ta \tand\\
        % \quad \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]},\ms_S,\ms_T \ldotp\\
        % \qquad\pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        % \qquad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        % \qquad \decInstr{\mscode([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        % \qquad \mscode(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets
\end{multline*}

Unlike return seals, closure seals are allowed in both trusted and untrusted components, but we cannot allow them to use closure seals to seal the same things.
If we allowed this, then a trusted component could seal a read-only capability for its code (i.e.\ access to return seals) with an untrusted closure seal and give it to the untrusted component which can unseal it with \texttt{xjmp}.
The closure seals in untrusted code should therefore only be used to seal things from the untrusted value relation, i.e.\ things that are safe to transfer to an untrusted component.
The trusted closure seals can be used to seal things from the untrusted value relation.
This allows trusted seals to seal capabilities for the trusted code which is necessary in order to create a new closure.

@@

Register files are related either by a trusted or untrusted relation $\mathcal{R}_\trust(W)$, and they are related when they contain words related in the appropriate value relation.
The register-file relation splits the ownership of $W$ between the registers as linear capability need a spatially owned region; but to prevent aliasing, the they cannot depend on the same spatially owned region.
% memory relation
The memory on \srccm{} consists of a heap $\ms_S$, a call stack $\stk$ and a free stack $\ms_\stk$.
This relates to a memory $\ms_T$ on \trgcm{} in a world $W$ if we can split $\ms_T$ into three parts that correspond with the view \srccm{} provides on the memory.
Each of the three partitions of $\ms_T$ must be related to one of the three parts of \srccm{} memory under the relevant sub-world of $W$.
The heap $\ms_S$ relates to the heap partition of $\ms_T$ in the standard way, i.e.\ the two memories can be partitioned, so they satisfy all the regions of the heap world.
The free stack $\ms_\stk$ also relates to the free stack partition of $\ms_T$ in the standard way but with respect to the free stack world.
%Further, the address $\stkb$ should always be an address of the free stack (the return token always has authority over $\stkb$), so it must be one of the addresses of $\ms_\stk$.
The call stack $\stk$ relates to the call stack partition of $\ms_T$ if it can be split into parts related to each of the stack frames.
Further, as each of the stack frames corresponds to a slot in the return order none of the stack frames should be empty.
It is the physical location of a stack frame that determines its order in the call stack, so the order must be preserved by the partitions of $\ms_T$.

@@

\subsubsection{Value relations}
The purpose of the value relations is twofold: It relates the words on \srccm{} to the words they correspond to on \trgcm{}, and it defines what pairs of words are safe, i.e.\ are unable to break the guarantees of the capability machines including the LSE and WBCF guaranteed by \srccm{}.
As previously explained, we try to ensure LSE and WBCF for some trusted code that we expect to behave reasonably, e.g.\ by handling return seals in a safe way.
This means that it is safe to give return seals to trusted code; but at the same time, we have some code that we do not trust and want to guarantee LSE and WBCF against.
This means that there is a difference between what is safe to give to the trusted code and the untrusted code which we capture by having a trusted and an untrusted value relation.
The two value relations are sketched in Figure~\ref{fig:value-relation}.

The untrusted value relation relates words from the two machines that can safely be given to untrusted code.
Data is always safe, and a piece of data relates to itself.
A stack pointer is related to a linear capability with the same (non-executable) permission, range of authority, and current address.
A safe stack pointer with read permission should only be able to produce safe words, i.e.\ words in $\mathcal{V}_\untrusted$, and a safe stack pointer with write permission can at least write safe words.
Regions model the contents of memory, so to determine safety of a stack pointer there must be an appropriate spatially owned region in the free stack sub-world.
Syntactically equal memory capabilities are related in a similar way to stack capabilities when they have read (and write) permission.
The difference is that memory capabilities point to the heap rather than the free stack, so the regions they depend on are picked from the heap sub-world.
An executable memory capability should always point to code which means that there should be a code region governing the memory it references.
The trusted return seals and closure seals cannot be given safely to untrusted code, so a set of seals is only in the untrusted value relation when it is disjoint from those, and they may only be used to seal things in $\mathcal{V}_\untrusted$.
% The write condition is called $\stackWriteCond{}$
% \begin{multline*}
%   \stackWriteCond{W} =\\
%   \left\{
%     \begin{array}{l}\npair{A} \mid 
%       \exists S \subseteq \addressable{\linear, \pwfree} \ldotp  \exists R : S \fun \powerset{\nats} \\
%       \quad \biguplus_{r\in S} R(r) \supseteq A \tand \forall r \in S \ldotp |R(r)| = 1  \tand \\
%       \quad \forall r \in S \ldotp \pwfree(r).H \nsupeq \stdreg{R(r),\gc}{\pur}.H \tand \pwfree(r) \text{ is address-stratified}
%     \end{array}
%   \right\}
% \end{multline*}
When a piece of untrusted code is called, it is passed a return pair consisting of two sealed capabilities.
Calling untrusted code should be safe, so sealed capabilities must be related in the untrusted value relation.
The seal interpretation function decides what a seal can be used for, so for a pair $(\src{\sealed{\sigma,\vsc_S}}, \sealed{\sigma,\vsc_T})$ to be in the untrusted value relation, $(\src{\vsc_S},\vsc_T)$ must be in the seal interpretation function for $\sigma$.
Further, it should be safe to use the sealed capability, i.e.\ jump to it with \texttt{xjmp}, which is exactly what $\mathcal{E}_{\mathrm{xjmp}}$ expresses, so when paired up with another pair from the seal interpretation function for $\sigma$ the quadruple must be in $\mathcal{E}_{\mathrm{xjmp}}$.

@@

The logical relation presented in Section~\ref{subsec:logical-relation} is lifted to a component relation $\mathcal{C}$ (we will also use $\cong$ to denote this relation) and an executable configuration relation $\mathcal{EC}$ in a straight forward manner.

@@

    $\dom(\comp.\mscode) \subseteq \ta$ and $\src{\comp}$ is a reasonable component; or
